---
title: "MIKE worksheet"
output: html_notebook
---

```{r}
library(tidyverse)
library(RColorBrewer)
library(lubridate)
theme_set(theme_bw() + theme(text = element_text(size=14), 
                             panel.border = element_blank(), panel.grid.major = element_blank(),
                             panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))) 
```


#TODO: Trend in cause/means over time
#TODO: create cause-specific map 
#TODO: ratio of conflict events to conflict killings within each MEP HWC sector

```{r}
## Load Data
mike <- read.csv("C:/Users/nhahn/Dropbox (Personal)/MEP_Working/Data/ER_Events/events_csv/MEP-MIKE.csv")

# time formatting
mike$time <- as.POSIXct(strptime(mike$time, format = '%Y-%m-%d %H:%M:%S'))

# split CauseOfDeath into cause and means
mike <- separate(mike,event_details_CauseOfDeath, c("cause", "means"), sep = "_", remove = FALSE, extra = "merge")
mike$cause <- as.factor(mike$cause)
mike$means <- as.factor(mike$means)

```

### Chi Squared Test
Test if type of illegal killing (poached vs. conflict) is correlated with means of killing
1. filter to conflict and poaching events
2. create contingency table
3. run chi2 test

```{r}
tbl <- droplevels(filter(mike, cause %in% c("conflict", "poached")))
tbl <- table(tbl$means, tbl$cause)

(tbl)
round(prop.table(tbl),2)
```

pvalue is << 0.05. Cause and Means are highly correlated. This is likely due to the large number of spear poaching events -- 50% of all poaching events in the data set. However, means does not appear automatically associated with cause, as there are conflict deaths from gunshot and spear. 
```{r}
chi2 <- chisq.test(tbl)
chi2
```

Line graph of Means by year carcass found. Seperate graphs for Cause. 
```{r}
df <- droplevels(filter(mike, cause %in% c("conflict", "poached"))) %>%
  group_by(year(time), cause) %>%
  tally() %>%
  mutate(freq = n/sum(n)) %>%
  rename(year = `year(time)`)
ggplot(df, aes(x = as.factor(year), y = freq, group = cause, shape = cause)) + geom_point() + geom_line() +
  labs(x = "Year", y = "Frequency (By Year)", shape = "Type of Death")

df <- droplevels(filter(mike, cause %in% c("conflict", "poached"))) %>%
  group_by(year(time), means) %>%
  tally() %>%
  mutate(freq = n/sum(n)) %>%
  rename(year = `year(time)`)
ggplot(df, aes(x = as.factor(year), y = freq, group = means, shape = means)) + geom_line() + geom_point() + scale_shape_manual(values = as.integer(unique(df$means)))


```


